A Comparison of Graph Processing Systems

### 2020-09-24 documentation/paper/main.tex
Rückmeldung übernommen und todos eingefügt

\begin{abstract}
Graphs are growing fast and are becoming increasingly popular in many different domains.
There are several different frameworks that make it easy to write graph algorithms.
They often use different approaches and have significant performance differences on different graphs.
In this paper we compare some state-of-the-art non-uniform memory access (NUMA) aware systems in terms of performance, both on real world and synthetic graphs.
Therefore, we consider the three algorithms Single-Source Shortest-Paths (SSSP), Breadth-First Search (BFS) and PageRank (PR).
We compare not only the time the frameworks indicate, they need to compute, but also the time they actually run and consume resources, and the resulting overhead time.
\end{abstract}

>>>>

\begin{abstract}
Graphs are growing fast and are becoming increasingly popular in many different domains, like \todo{...}
There are several different frameworks that help writing graph algorithms.
They often use different approaches and have significant performance differences on different graphs.
\todo{Giraph nicht NUMA aware, da unten raus nehmen..}
In this paper we compare some state-of-the-art non-uniform memory access (NUMA) aware systems in terms of performance, both on real world and synthetic graphs.
\todo{umformulieren}
Therefore, we consider the three algorithms: Single-Source Shortest-Paths (SSSP), Breadth-First Search (BFS) and PageRank (PR).
We compare not only the time the frameworks indicate, they need to compute, but also the time they actually run and consume resources, and the resulting overhead time.
\end{abstract}

### 2020-09-24 documentation/paper/sections/conclusion.tex
Rückmeldung übernommen und todos eingefügt

We compared the performance of several graph processing frameworks including Ligra, Polymer, Gemini, Galois and Giraph. 
For comparison we used the three problems SSSP, BFS and PR on graphs of various sizes.
We measured the calculation time of the algorithm itself and the overall execution time, which also includes the startup time of the programm.
All of the frameworks were tested on a single compute node. Gemini, Galois and Giraph also support distributed computation and were therefore also tested on a five-node cluster.

On a single node, the fastest framework in computation time is dependent on the problem under consideration.
For SSSP, we found Polymer to perform best while for BFS Ligra and Gemini are fastest. PR was calculated by Galois the fastest.
When looking at the computation times of the distributed systems, it stands out that the best time for each problem is beaten by most if not all of the single node times.
Comparing the distributed computation times SSSP and BFS results looked similar with Giraph being the fastest with the exception of synthetic graphs where Gemini was faster. The push version of Galois outperforms the pull version but ultimatively was slower than Gemini. For distributed PR, Gemini was the fastest by a margin.
In contrast, in terms of overall execution times, Galois almost always outperforms the other frameworks because the overhead is very small.

>>>>

We compared the performance of several graph processing frameworks including Ligra, Polymer, Gemini, Galois and Giraph. 
For comparison we used the three graph algorithms SSSP, BFS and PR on data sets of various sizes.
We measured the calculation time of the algorithm itself and the overall execution time, which also includes the framework overhead.
All of the frameworks were tested on a single compute node. Gemini, Galois and Giraph also support distributed computation and were therefore also tested on a five-node cluster.

\todo{Ab hier kürzer fassen}
On a single node, the fastest framework in computation time is dependent on the problem under consideration.
For SSSP, we found Polymer to perform best while for BFS Ligra and Gemini are fastest. PR was calculated by Galois the fastest.
When looking at the computation times of the distributed systems, it stands out that the best time for each problem is beaten by most if not all of the single node times.
Comparing the distributed computation times SSSP and BFS results looked similar with Giraph being the fastest with the exception of synthetic graphs where Gemini was faster. The push version of Galois outperforms the pull version but ultimatively was slower than Gemini. For distributed PR, Gemini was the fastest by a margin.
In contrast, in terms of overall execution times, Galois almost always outperforms the other frameworks because the overhead is very small.

### 2020-09-24 documentation/paper/sections/discussion.tex
Rückmeldung übernommen und todos eingefügt

%## production
On production systems, we expect it to be possible to determine what configuration is best before putting the system in service.
But not only the absolute runtime of each computation has to be taken into account. Production systems have to be very reliable because downtime often directly correlates to financial loss.
This is often a reason for choosing a distributed system. Many nodes together are less prone to failure of the entire system.
Thus, there is an argument to be made for systems like Hadoop.
It automatically handles node failures, with a node failure not immediately resulting in loss of data or faulty computation results.
This is especially important on very large clusters, that can not easily be monitored by humans.
Further, a user defining an application does not need to know the topology of the system that runs the application later on.
Hadoop transparently splits the input data and distributes work among the worker nodes.
Additionally, Giraph using hadoop shows a bias towards larger graphs, we observed an improvement in performance over single-node Giraph, when increasing the size of the input graph.
This shows the hadoop infrastructure to be working well with large graphs, which is exactly what the distributed system should be used for. 
Only a system like hadoop can reasonably be used on such large clusters.
Any of the other framewors will just result in large administration overhead due to node failures or other problems of distributed systems in general.
Furthermore, the calculation times of Giraph proved to be very fast in comparison to the other distributed frameworks, with Giraph often being one of the fastest framework in computation time.
Very short calculation times along with automatic handling of the computation cluster make Giraph using Hadoop a good choice for production systems.
Pregel is a similar, fault-tolerant and scalable system but it is contrary to Giraph closed-source.

But not only distributed production systems are possible, of course. 
For small graphs, i.e. those that are not too large for single node RAM, a single-node setup is usually faster than the same framework on a distributed cluster. 
We have shown that the distributed cluster of the same framework becomes faster than the equivalent single-node setup only on large graphs.
We have found around 2B edges, or roughly the size of the twitter graph to be a size at which Gemini and Giraph are faster on the distributed setup compared to single-node.
If required, horizontal scaling can for example be used to provide fault-tolerance of the entire system.

If only the calculation time is important, there are multiple options with the deciding factors being the graph size and the application.
For algorithms that perform a traversal through the graph, with a concentrated and relatively small number of active vertices, our results for SSSP and BFS are most relevant.
Here Galois is fastest on very small graphs (i.e. flickr), but is quickly overtaken by other frameworks.
In the case of SSSP those are Gemini and Polymer and on BFS they are Ligra and again Gemini. While Gemini is not the fastest on either BFS or SSSP, it is still the safest bet on other algorithms. That is because Polymer is fast on SSSP but slowest on BFS and in turn Ligra is fastest on BFS but one of the slowest frameworks on SSSP. Gemini is second fastest on SSSP and BFS on the larger graphs.
Also, it might be important to consider the implementation style of a custom application for either of the frameworks.
Applications similar to BFS and SSSP are expected to perform better with a push style implementation. 
This is because they are algorithms with only few active vertices or where the affected push areas are not overlapping. 
Hence, only little synchronization is needed that could be impacting performance negatively.

>>>>

%## production
On production systems, we expect it to be possible to determine what configuration is best before putting the system in service.
But not only the absolute runtime of each computation has to be taken into account. Production systems have to be very reliable because downtime often directly correlates to financial loss.
This is often a reason for choosing a distributed system. Many nodes together are less prone to failure of the entire system.
Thus, there is an argument to be made for systems like Hadoop.
It automatically handles node failures, with a node failure not immediately resulting in loss of data or faulty computation results.
This is especially important on very large clusters, that can not easily be monitored by humans.
Further, a user defining an application does not need to know the topology of the system that runs the application later on.
Hadoop splits the input data and distributes work among the worker nodes.
Additionally, Giraph using hadoop shows a bias towards larger graphs. We observed an improvement in performance over single-node Giraph, when increasing the size of the input graph.
This shows the hadoop infrastructure to be working well with large graphs, which is exactly what the distributed system should be used for. 
Only a system like hadoop can reasonably be used on such large clusters.
Any of the other evaluated frameworks will just result in large administration overhead due to node failures or other problems of distributed systems in general.
Furthermore, the calculation times of Giraph proved to be very fast in comparison to the other distributed frameworks, with Giraph often being one of the fastest framework in computation time.
Very short calculation times along with automatic handling of the computation cluster make Giraph using Hadoop a good choice for production systems.
GraphX and Pregel are similar, fault-tolerant and scalable systems. GraphX uses Apache Spark, extending on hadoop. Pregel is contrary to Giraph or GraphX closed-source.

But not only distributed production systems are possible, of course. 
For small graphs, i.e. those that are not too large for single node RAM, a single-node setup is usually faster than the same framework on a distributed cluster. 
We have shown that the distributed cluster of the same framework becomes faster than the equivalent single-node setup only on large graphs.
We have found that at around 2B edges (roughly the size of the twitter graph) Gemini and Giraph are faster on the distributed setup compared to single-node.
If required, horizontal scaling can for example be used to provide fault-tolerance of the entire system.

If only the calculation time is important, there are multiple options with the deciding factors being the graph size and the application.
For algorithms that perform a traversal through the graph, with a concentrated and relatively small number of active vertices, our results for SSSP and BFS are most relevant.
Here Galois is fastest on very small graphs (i.e. flickr), but is quickly overtaken by other frameworks.
In the case of SSSP those are Gemini and Polymer and on BFS they are Ligra and again Gemini. While Gemini is not the fastest on either BFS or SSSP, it is still a good choice on other algorithms. That is because Polymer is fast on SSSP but slowest on BFS and in turn Ligra is fastest on BFS but one of the slowest frameworks on SSSP. Gemini is second fastest on SSSP and BFS on the larger graphs.
Also, it might be important to consider the implementation style of a custom application for either of the frameworks.
Applications similar to BFS and SSSP are expected to perform better with a push style implementation. 
This is because they are algorithms with only few active vertices or where the affected push areas are not overlapping. 
Hence, only little synchronization is needed that could be impacting performance negatively.

### 2020-09-24 documentation/paper/sections/introduction.tex
some more minor fixes
Rückmeldung übernommen und todos eingefügt

%!TEX root=../main.tex
\section{Introduction}

In recent years, graph processing has become increasingly important, both in academic and industrial environments \cite{Gemini}.
In many areas, such as machine learning and data mining, but also scientific computations, problems can be modeled and solved with the help of graphs \cite{Polymer}.
Thus a new application domain called graph analytics was created.
Some business models are based on working with graphs, such as marketing analysis, advertising, or even Google's search engine \cite{pagerank}.
In addition, the size of graphs has grown significantly in the last few years and will most likely continue to grow.
This makes the performance and memory efficency of graph analysis applications more important than ever.
But also the parallelism and distribution of graph algorithms becomes more important with increasing graph sizes.

Some frameworks were developed for this purpose.
One of the most commonly known graph-processing systems is the closed-source and distributed system Pregel \cite{pregel}.
Pregel, as well as some open source versions of it, like Giraph \cite{Giraph} and GraphX \cite{graphx}, was built to handle large graphs reliably and be fault-tolerant on large MapReduce infrastructures.
Due to the lack of cache locality, as these systems are written object-oriented, they provide rather suboptimal performance.
Because of this, many non-uniform memory access (NUMA) aware systems such as Ligra \cite{Ligra}, Polymer \cite{Polymer}, Galois \cite{Galois} and Gemini \cite{Gemini} were introduced.
These systems have some similarities, but many differences. This results in considerable performance differences on many occasions.
The systems were initially compared at their introduction.
Because most of the systems are several years old and some of them have evolved, or are still evolving, these comparisons were often incomplete or are simply no longer relevant today. 
Other systems are simply no longer used today, which also reduces the significance of some of the existing comparisons.

@@

The rest of this paper is organized as follows.
In section 2 we first clarify what a graph is, how the algorithms we are looking at work and what else is necessary for further understanding. Then we give a short overview of the frameworks we comparing in section 3 and of the graph formats used in section 4. 
Before we come to our results, we show in section 5 what has been done in this area so far. 
In Section 6, which is the core of the paper, we will present all our results and also briefly show how they were developed. 
After that, in section 7, we will discuss these results. And at the end in section 8 we will present our conclusion.

>>>>

%!TEX root=../main.tex
\section{Introduction}
\todo{Hier gibts einige Todos, am besten nochmal nachschauen}
In recent years, graph processing has become increasingly important, both in academic and industrial environments \cite{Gemini}.
In many areas, such as machine learning and data mining, but also scientific computations, problems can be modeled and solved with the help of graphs \cite{Polymer}.
Thus a new application domain called graph analytics was created.
Some business models are based on working with graphs, such as marketing analysis, advertising, or even Google's search engine \cite{pagerank}.
In addition, the size of graphs has grown significantly in the last few years and will most likely continue to grow.
This makes the performance and memory efficency of graph analysis applications more important than ever.
But also the parallelism and distribution of graph algorithms becomes more important with increasing graph sizes.

Some frameworks were developed for this purpose.
One of the most commonly known graph-processing systems is the closed-source and distributed system Pregel \cite{pregel}.
Pregel, as well as some open source versions of it, like Giraph \cite{Giraph} and GraphX \cite{graphx}, was built to handle large graphs reliably and be fault-tolerant on large MapReduce infrastructures.
Due to the lack of cache locality, as these systems are written object-oriented, they provide rather suboptimal performance.
Because of this, many non-uniform memory access (NUMA) aware systems such as Ligra \cite{Ligra}, Polymer \cite{Polymer}, Galois \cite{Galois} and Gemini \cite{Gemini} were introduced.
These systems have some similarities, but many differences. This results in considerable performance differences on many occasions.
The systems were initially compared at their introduction.
Because most of the systems are several years old and some of them have evolved, or are still evolving, these comparisons were often incomplete or are simply no longer relevant today. 
Other systems are no longer in use, which also reduces the significance of some of the existing comparisons.

@@

The rest of this paper is organized as follows:
Section 2 is dedicated to all preliminaries, in order to keep this paper self-contained.
After that, we give a short overview of each framework in our comparison.
Section 4 explains the used graph formats. 
Related work is discussed in section 5.
In Section 6 we present our evaluation, followed by a discussion of these results in section 7. Finally we conclude the paper in section 8.

### 2020-09-24 documentation/paper/sections/overview/giraph.tex
various fixes
Rückmeldung übernommen und todos eingefügt

In order for Giraph a achieve scalability and parallelization, it is built on top of Apache Hadoop \cite{Giraph}.
Hadoop is a MapReduce infrastructure providing a fault tolerant basis for large scale data processing.
Hadoop supplies a distributed file system (HDFS), on which all computations are performed.
Giraph is thus, even when only using a single node, running in a distributed manner.
Hence, expanding single-node processing to a multi-node cluster is seamless.
Giraph uses the Map functionality of Hadoop to run the algorithms. Reduce is only used as the identity function.

>>>>

In order for Giraph to achieve scalability and parallelization, it is built on top of Apache Hadoop \cite{Giraph}.
Hadoop is a MapReduce infrastructure providing a fault tolerant basis for large scale data processing.
Hadoop supplies a distributed file system (HDFS), on which all computations are performed.
Giraph is thus, even when only using a single computation node, running in a distributed manner.
Hence, expanding single-node processing to a multi-node cluster is seamless.
Giraph uses the Map functionality of Hadoop to run the algorithms. Reduce is only used as the identity function.

### 2020-09-24 documentation/paper/sections/overview/overview.tex
merge und besprechung
Rückmeldung übernommen und todos eingefügt

\section{Framework Overview}
We start with a short overview describing functionality and characteristics of several state-of-the-art graph processing frameworks.
All of the following frameworks are part of our testing.

>>>>

\section{Framework Overview}
The following paragraphs present a short overview describing functionality and characteristics of several state-of-the-art graph processing frameworks.
All of the described frameworks are part of our evaluation.

### 2020-09-24 documentation/paper/sections/preliminiaries.tex
Rückmeldung übernommen und todos eingefügt

\subsection{Graphs and Paths}
An \emph{unweighted graph} is the pair $G=(V,E)$ where the \emph{vertex set} is $V\subseteq\mathbb N$ and the $E$ is the \emph{edge set}.
The edge set describes a number of connections or relations between two vertices. Depending on these relations, a graph can be directed or undirected. For a \emph{directed graph} the edge set is defined as
\begin{equation*}
  E\subseteq\{(x,y)\,|\, x,y\in V, x\neq y\}
\end{equation*}
and in the \emph{undirected} case, the direction is no longer relevant. Thus, in an undirected graph for each $(x,y)\in E$, it holds $(x,y)=(y,x)$.
The size of a graph is defined as the number of edges $|E|$ \cite{newman2010networks}.
Independently of the graph being directed or not, a graph can be \emph{weighted}. In this case a function $w:E\rightarrow \mathbb R$ is introduced, that maps an edge to a numerical value, further describing the relation.

@@

\subsection{Breadth-First Search}
Breadth-first search (BFS) is a search problem on a graph.
It requires an unweighted graph and a start vertex as input.
The output is a set of vertices that are reachable from the start vertex.
In some special cases, a target vertex is also given. In the case of a target being given, the output is true if a path from start to target exists and otherwise false.
It is called Breadth-First search because the algorithm searches in a path length-based way. First all paths of length one i.e. all neighbors of the start vertex are checked before checking paths of length two and so on.
The search algorithm, where the paths of maximum length are checked first is called Depth-First search.

\subsection{PageRank}
\emph{PageRank} (PR) is a link analysis algorithm that weighs the vertices of a graph, measuring the vertices relative importance within the graph \cite{pagerank}.
The algorithm was invented by Sergey Brin and Larry Page, the founders of Google. To this date, Google Search uses PageRank to rank web pages in their search engine results.
It is also used as a centrality metric.

The output of PageRank is a percentage for each vertex. This percentage, called the PageRank of the vertex, is the probability of arriving at this vertex when starting at a random vertex and then following random edges.
With a small probability, the damping factor, a jump is made to a completely random vertex instead of following a edge.
The PageRank can be calculated by iterating
\begin{equation*}
	\operatorname{PR}(v_i) = \frac{1 - d}{|V|} + d \sum_{v_j \in \operatorname{N^-}(v_i)} \frac{\operatorname{PR}(v_j)}{\operatorname{deg^+}(v_j)}
\end{equation*}
for all edges $v_i \in V$.
$d$ is the damping factor, $\operatorname{deg^+}(v)$ is the number of outgoing edges the vertex $v$ has and $\operatorname{N^-}(v)$ denotes the set of vertices that have edges to vertex $v$.

@@

\subsection{Push and Pull Variants}
Many parallel graph algorithms, including the three we consider here, are implemented by iterating over vertices \cite{tao}.
The vertex, to which the operator is applied, is called \emph{active}.
It's possible to iterate over edges and consider them as \emph{active}, but in the following, whithout loss of generality, we will use the term \emph{active vertex}.
This operator considers only a so called \emph{neighborhood} to the \emph{active} vertex.
To this \emph{neighborhood} belong only vertices, which are in the direct surrounding of the \emph{active} vertex.
This allows to parallelise the iteration relatively easy, caused by the locality of the operator.

@@

A user specifies just the two functions Map and Reduce.
The system hides the details of parallelization, fault-tolerance, data distribution and load balancing away from the application logic.
All of these features are automatically provided.
Execution is performed in three phases:
\begin{enumerate}
	\item Map phase: The input data is distributed between a set of Map processes, the Map functionality is specified by the user. Ideally all Map processes run in parallel so the map processes need to be independent. Results from this phase are written into (multiple) intermediate storage points.
	\item Shuffle phase: The results are grouped according to a key provided by the Map algorithm. Each set of results is then handed to one system for the next phase.
	\item Reduce phase: Every set of intermediate results is input to exactly one reduce process. The Reduce functionality is again specified by the user and ideally runs in parallel.
\end{enumerate}
Giraph \cite{Giraph} is an example of a system using this framework.

@@

Translating from the virtual to the physical address space significantly reduces the performance of today's data processing systems, since the size of the RAM grows much faster than the size of the TBL \cite{hugepages, superpages}.
To counteract this problem, hugepages were developed, which are usually several mibibibytes in size.
This minimizes the CPU time needed for table lookups, because there are fewer TBL misses.
This speedup is especially noticeable in very memory intensive applications, like graph processing systems.

>>>>

\subsection{Graphs and Paths}
An \emph{unweighted graph} is the pair $G=(V,E)$ where the \emph{vertex set} is $V\subseteq\mathbb N$ and the $E$ is the \emph{edge set}.
The edge set describes a number of connections or relations between two vertices. Depending on these relations, a graph can be directed or undirected. For a \emph{directed graph} the edge set is defined as
\begin{equation*}
  E\subseteq\{(x,y)\,|\, x,y\in V, x\neq y\}
\end{equation*}
and in the \emph{undirected} case, the direction is no longer relevant. Thus, in an undirected graph for each $(x,y)\in E: (x,y)=(y,x)$ holds.
The size of a graph is defined as the number of edges $|E|$ \cite{newman2010networks}.
Independently of the graph being directed or not, a graph can be \emph{weighted}. In this case a function $w:E\rightarrow \mathbb R$ is introduced, that maps an edge to a numerical value, further describing the relation.

@@

\subsection{Breadth-First Search}
Breadth-first search (BFS) is a search problem on a graph.
It requires an unweighted graph and a start vertex as input.
The output is a set of vertices that are reachable from the start vertex.
In some special cases, a target vertex is also given. In the case of a target being given, the output is true if a path from start to target exists and otherwise false.
It is called Breadth-First search because the algorithm searches in a path length-based way. First all paths of length one i.e. all neighbors of the start vertex are checked before checking paths of length two and so on.

\subsection{PageRank}
\emph{PageRank} (PR) is a link analysis algorithm that weighs the vertices of a graph, measuring the vertices relative importance within the graph \cite{pagerank}.
The algorithm was invented by Sergey Brin and Larry Page, the founders of Google. To this date, Google Search uses a modified version of PageRank to rank web pages in their search engine results.
It is also used as a centrality metric.

The output of PageRank is a percentage for each vertex. This percentage, called the PageRank of the vertex, is the probability of arriving at this vertex when starting at a random vertex and then following random edges.
With a small probability, the damping factor, a jump is made to a completely random vertex instead of following an edge.
The PageRank can be calculated by iterating
\begin{equation*}
	\operatorname{PR}(v_i) = \frac{1 - d}{|V|} + d \sum_{v_j \in \operatorname{N^-}(v_i)} \frac{\operatorname{PR}(v_j)}{\operatorname{deg^+}(v_j)}
\end{equation*}
for all edges $v_i \in V$.
$d$ is the damping factor, $\operatorname{deg^+}(v)$ is the number of outgoing edges the vertex $v$ has and $\operatorname{N^-}(v)$ denotes the set of vertices that have edges to vertex $v$.

@@

\subsection{Push and Pull Variants}
\todo{kompliziert}
Many parallel graph algorithms, including the three we consider here, are implemented by iterating over vertices \cite{tao}.
The vertex, to which the operator is applied, is called \emph{active}.
It's possible to iterate over edges and consider them as \emph{active}, but in the following, whithout loss of generality, we will use the term \emph{active vertex}.
This operator considers only a so called \emph{neighborhood} to the \emph{active} vertex.
To this \emph{neighborhood} belong only vertices, which are in the direct surrounding of the \emph{active} vertex.
This allows to parallelise the iteration relatively easy, caused by the locality of the operator.

@@

A user specifies just the two functions Map and Reduce.
The system hides the details of parallelization: fault-tolerance, data distribution and load balancing away from the application logic.
All of these features are automatically provided.
Execution is performed in three phases:
\begin{enumerate}
	\item Map phase: The input data is distributed between a set of Map processes, the Map functionality is specified by the user. Ideally all Map processes run in parallel so the map processes need to be independent. Results from this phase are written into (multiple) intermediate storage points.
	\item Shuffle phase: The results are grouped according to a key provided by the Map algorithm. Each set of results is then handed to one system for the next phase.
	\item Reduce phase: Every set of intermediate results is input to exactly one reduce process. The Reduce functionality is again specified by the user and ideally runs in parallel.
\end{enumerate}
Giraph \cite{Giraph} is an example of a system using this framework.

@@

Translating from the virtual to the physical address space significantly reduces the performance of today's data processing systems, since the size of the RAM grows much faster than the size of the TBL \cite{hugepages, superpages}.
To counteract this problem, hugepages were developed, which are usually several mibibytes in size.
This minimizes the CPU time needed for table lookups, because there are fewer TBL misses.
This speedup is especially noticeable in very memory intensive applications, like graph processing systems.

### 2020-09-24 documentation/paper/sections/results/bfs.tex
some work on discussion/comparison
merge und besprechung
various fixes
Rückmeldung übernommen und todos eingefügt

\todo{Einleitung, analog SSSP}

\paragraph{Single-Node}
\begin{figure*}
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_calcTime.png}
		\caption{Calculation time}
		\label{fig:singleNodeBFS_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_execTime.png}
		\caption{Execution time}
		\label{fig:singleNodeBFS_exec}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_overheadTime.png}
		\caption{Overhead time}
		\label{fig:singleNodeBFS_overheadNormalized}
	\end{subfigure}
	\caption{Average times for BFS on a single computation node, black bars represent one standard deviation in our testing}
\end{figure*}
Just like with SSSP, Giraph ran out of memory ($>$250 GB) on any graph larger than wikipedia. Also, Polymer failed to complete on rMat28. Ligra, that failed during SSSP however completed our benchmark for BFS.

@@

For the remaining frameworks, we compare to Ligra since it is generally slightly faster than Gemini.
Giraph and Polymer are one to two orders of magnitude slower than Ligra. The only exception to this is Polymer on flickr, here Polymer is just 24ms (35\%) slower than Ligra.
Giraph takes between 13$\times$ and 32$\times$ longer than Ligra on flickr, orkut or wikipedia.
As we said, Polymer is comparable to Ligra on flickr, but for the other graphs, its computation times are even longer than those of Giraph. The upside is though, that Polymer managed -- contrary to Giraph -- to finish computation on some larger graphs.
Polymer takes between 55$\times$ and 530$\times$ longer than Ligra on the graphs larger than flickr.
Especially interesting is here the difference between the times of rMat27 and friendster. There are 20\% (439M) more edges in friendster, yet the computation for the synthetic rMat27 takes 42\% longer.
Galois calculation performance is comparable to Gemini or Ligra on the smaller three graphs (flickr, orkut and wikipedia).
Only on the larger graphs is Galois slower than Gemini or Ligra, meanwhile Galois is always faster than Polymer by one order of magnitude.

The execution time results are very similiar to our findings of SSSP (cf. \autoref{fig:singleNodeBFS_exec}). Just like on SSSP, Galois is the fastest framework and Ligra is second fastest on all graphs except flickr.
Again, Gemini and Polymer are comparable in their performance and at the same time the slowest frameworks on all graphs except flickr, orkut and wikipedia.
On those three graphs, Giraph is the slowest, by a difference of one to two orders of magnitude compared to Galois.
On the other graphs, Polymer and Gemini are one order of magnitude slower than Galois.

\paragraph{Distributed}
\begin{figure*}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedBFS_calcTime.png}
		\caption{Calculation time}
		\label{fig:distributedBFS_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedBFS_execTime.png}
		\caption{Execution time}
		\label{fig:distributedBFS_exec}
	\end{subfigure}
	\hfil
	% \begin{subfigure}{0.3\textwidth}
	% 	\includegraphics[width=\linewidth]{../../plots/distributedBFS_overheadTimeNormalized.png}
	% 	\caption{Overhead time normalized by the graph size in million edges}
	% 	\label{fig:distributedBFS_overheadNormalized}
	% \end{subfigure}
	% \hfil
	\caption{Average times for BFS on the distributed cluster, black bars represent one standard deviation in our testing}
	\label{fig:distributedBFS}
\end{figure*}
For both the calculation and the execution times, Breadth-First Search shows similar behaviour as the distributed SSSP test case. This is expected since both are graph traversal algorithms starting in one source vertex.
Hence, calculation complexity for each vertex and communication overhead is similar. All measurements can be seen in \autoref{fig:distributedBFS}.

@@

\todo{Discussion ausarbeiten}

>>>>

\paragraph{Single-Node}
\begin{figure*}
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_calcTime.png}
		\caption{Calculation time}
		\label{fig:singleNodeBFS_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_execTime.png}
		\caption{Execution time}
		\label{fig:singleNodeBFS_exec}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_overheadTime.png}
		\caption{Overhead time}
		\label{fig:singleNodeBFS_overheadNormalized}
	\end{subfigure}
	\caption{Average times for BFS on a single computation node}
\end{figure*}
Giraph ran out of memory ($>$250 GB) on any graph larger than wikipedia. Also, Polymer failed to complete on rMat28. Ligra, that failed during SSSP however completed our benchmark for BFS.

@@

We compare the remaining frameworks to Ligra because it is generally slightly faster than Gemini.
Giraph and Polymer are one to two orders of magnitude slower. The only exception to this is Polymer on flickr, here Polymer is just 24ms (35\%) slower than Ligra.
Giraph takes between 13$\times$ and 32$\times$ longer than Ligra on flickr, orkut or wikipedia.
As already stated, Polymer is comparable to Ligra on flickr, but for the other graphs, its calculation times are even longer than those of Giraph. However, Polymer managed -- contrary to Giraph -- to finish computation on some larger graphs.
Polymer takes between 55$\times$ and 530$\times$ longer than Ligra on the graphs larger than flickr.
Here, the difference between the times of rMat27 and friendster are especially interesting. There are 20\% (439M) more edges in friendster, yet the computation for the synthetic rMat27 takes 42\% longer.
Galois calculation performance is comparable to Gemini or Ligra on the smaller three graphs (flickr, orkut and wikipedia).
Only on the larger graphs is Galois slower than Gemini or Ligra, meanwhile Galois is always faster than Polymer by one order of magnitude.

The execution time results are very similiar to our findings of SSSP (cf. \autoref{fig:singleNodeSSSP_exec}). Just like on SSSP, Galois is the fastest framework and Ligra is second fastest on all graphs except flickr.
Again, Gemini and Polymer are comparable in their performance and at the same time the slowest frameworks on all graphs except flickr, orkut and wikipedia.
On those three graphs, Giraph is the slowest, by a difference of one to two orders of magnitude compared to Galois.
On the other graphs, Polymer and Gemini are one order of magnitude slower than Galois.

\paragraph{Distributed}
\begin{figure*}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedBFS_calcTime.png}
		\caption{Calculation time}
		\label{fig:distributedBFS_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedBFS_execTime.png}
		\caption{Execution time}
		\label{fig:distributedBFS_exec}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedBFS_overheadTime.png}
		\caption{Overhead}
		\label{fig:distributedBFS_overhead}
	\end{subfigure}
	\hfil
	\caption{Average times for BFS on the distributed cluster}
	\label{fig:distributedBFS}
\end{figure*}
For both the calculation and the execution times, Breadth-First Search shows similar behaviour as the distributed SSSP test case. This is expected since both are graph traversal algorithms starting in one source vertex.
Hence, calculation complexity for each vertex and communication overhead is comparable. All measurements can be seen in \autoref{fig:distributedBFS}.

@@
